Synthetic algebraic geometry and the phenomenon of nongeometric sequents

### 2018-04-06 slides-pssl103.tex
Minor improvements

\begin{frame}{Curious universes}
  \begin{changemargin}{-1.1em}{0em}
    \small
    \begin{itemize}
      \item \intex{
        $\Eff \models \text{``There are infinitely many prime numbers.''}$
      }{\textcolor{green!90}{\cmark}}{
        External meaning: There is a Turing \hil{machine} producing arbitrarily many prime
        numbers.
      }

@@

      \item \intex{
        $\Sh(X) \not\models \text{``Any cont. function with opposite signs has
        a zero.''}$
      }{\textcolor{red!80}{\xmark}}{
        External meaning: Zeros can be picked \hil{locally continuously} in
        continuous families of continuous functions.
      }
    \end{itemize}
  \end{changemargin}

@@

  \begin{itemize}
    \item For an~$S$-scheme~$X$, its functor of points $\ull{X} =
    \Hom_S(\cdot,X)$ is an object of~$\Zar(S)$. It feels like \hil{the set of
    points} of~$X$.
    \item In particular, there is the ring object $\affl$.
    \item $\Zar(S)$ classifies local~$\O_S$-algebras which are local
    over~$\O_S$.
  \end{itemize}
\end{frame}

@@

  $\hil{$\boldsymbol{TX}$} = \Hom(\Delta, X)$, where $\Delta = \{ \varepsilon \? \affl \,|\, \varepsilon^2 = 0 \}$

@@

\begin{frame}{Synthetic quasicoherence}
  Recall~$\Spec(R) = \Hom_{\mathrm{Alg}(\affl)}(R, \affl)$ and consider the statement
  \[ \text{``the canonical map~$
    \begin{array}[t]{rcl}
      R &\longrightarrow& \Hom(\Spec(R), \affl) \\
      f &\longmapsto& (\alpha \mapsto \alpha(f))
    \end{array}
  $ is bijective''}.
  \]
  \vspace*{-1em}

@@

  \emph{Example.} Let~$f \? \affl$ such that~$f \neq 0$. Set~$R = \affl/(f)$.
  Then~$\Spec(R) = \emptyset$. Thus~$\Hom(\Spec(R),\affl)$ is a singleton.
  Hence~$R = 0$. Therefore~$f$ is invertible.
\end{frame}

>>>>

\begin{frame}{Curious universes}
  \begin{changemargin}{-1.1em}{0em}
    \small
    \begin{itemize}
      \item \intex{
        $\Eff \models \text{``There are infinitely many prime numbers.''}$
      }{\textcolor{green!90}{\cmark}}{
        External meaning: There is a \hil{Turing machine} producing arbitrarily many prime
        numbers.
      }

@@

      \item \intex{
        $\Sh(X) \not\models \text{``Any cont. function with opposite signs has
        a zero.''}$
      }{\textcolor{red!80}{\xmark}}{
        External meaning: Zeros can locally be picked \hil{continuously} in
        continuous families of continuous functions.
      }
    \end{itemize}
  \end{changemargin}

@@

  \begin{changemargin}{-1.0em}{-0.5em}
  \begin{itemize}
    \item For an~$S$-scheme~$X$, its functor of points $\ull{X} =
    \Hom_S(\cdot,X)$ is an object of~$\Zar(S)$. It feels like \hil{the set of
    points} of~$X$. \\[0.2em]
    \item \ \\[-1.2em]\mbox{In particular, there is the ring object $\affl$ with~$\affl(U) =
    \O_U(U)$.} \\[0.2em]
    \item $\Zar(S)$ classifies local~$\O_S$-algebras which are local
    over~$\O_S$.
  \end{itemize}
  \end{changemargin}
\end{frame}

@@

  $\hil{$\boldsymbol{TX}$} = X^\Delta$, where $\Delta = \{ \varepsilon \? \affl \,|\, \varepsilon^2 = 0 \}$

@@

\begin{frame}{Synthetic quasicoherence}
  Recall~$\Spec(R) = \Hom_{\mathrm{Alg}(\affl)}(R, \affl)$ and consider the statement
  \[ \text{``the canonical map~$
    \begin{array}[t]{rcl}
      R &\longrightarrow& (\affl)^{\Spec(R)} \\
      f &\longmapsto& (\alpha \mapsto \alpha(f))
    \end{array}
  $ is bijective''}.
  \]
  \vspace*{-1em}

@@

  \emph{Example.} Let~$x \? \affl$ such that~$x \neq 0$. Set~$R = \affl/(x)$.
  Then~$\Spec(R) = \emptyset$. Thus~$(\affl)^{\Spec(R)}$ is a singleton.
  Hence~$R = 0$. Therefore~$x$ is invertible.
\end{frame}
